How did the introduction of Stemmed Points affect Mobility and Site Occupation during the Late Pleistocene in Korea?

### 2021-12-09 analysis/paper/paper.Rmd
minor edits to respond to QSR manuscript checking

The entire R code [@Rlanguage2019] used for all the analysis and visualizations contained in this paper is included in the Supplementary Online Materials at https://doi.org/10.17605/OSF.IO/JXWY5 to enable re-use of materials and improve reproducibility and transparency [@Marwick2017]. Also in this version-controlled compendium [@marwick2018packaging] are the raw data for all the visualizations and tests reported here. All of the figures, tables, and statistical test results presented here can be independently reproduced with the code and data in this repository. The code is released under the MIT license, the data as CC-0, and figures as CC-BY, to enable maximum re-use.

@@

Among the sites in our sample there is a variation of about 5 degrees in the mean annual temperature (MAT), mostly controlled by elevation (Figure \@ref(fig:climate-model)). Through MIS 3, the temperature gradually decreased until the Last Glacial Maximum (LGM, 26.5-20 ka) (Clark et al., 2009). The MAT of MIS 2, including LGM, was relatively stable. The temperature increased again from late MIS 2 towards MIS 1. Compared to the east side of the Korean Peninsula, the west side tends to be relatively warmer. The range of MAT at the sites over our study period is 2-10℃. Gihwari and Gigok sites have the lowest (3℃) and Jeungsan-ri site has the highest MAT (8℃). The mean MAT for all sites fluctuated within 4℃ between 50 ka and 10 ka. The first appearance of stemmed points occurred in the middle of the decreasing MAT trend in MIS 3, at 40-35 ka. Figure 7D shows a negative relationship between temperature and elevation. For example, Gihwari Cave site, with one of the lowest MAT distributions, is located in the highest elevation around 300m above sea level. Jeungsanri is located at a lower elevation and has one of the highest MAT distributions.

>>>>

The entire R code [@Rlanguage2019] used for all the analysis and visualizations contained in this paper is openly available at https://doi.org/10.17605/OSF.IO/JXWY5 to enable re-use of materials and improve reproducibility and transparency [@Marwick2017]. Also in this version-controlled compendium [@marwick2018packaging] are the raw data for all the visualizations and tests reported here. All of the figures, tables, and statistical test results presented here can be independently reproduced with the code and data in this repository. The code is released under the MIT license, the data as CC-0, and figures as CC-BY, to enable maximum re-use.

@@

Among the sites in our sample there is a variation of about 5 degrees in the mean annual temperature (MAT), mostly controlled by elevation (Figure \@ref(fig:climate-model)). Through MIS 3, the temperature gradually decreased until the Last Glacial Maximum (LGM, 26.5-20 ka) (Clark et al., 2009). The MAT of MIS 2, including LGM, was relatively stable. The temperature increased again from late MIS 2 towards MIS 1. Compared to the east side of the Korean Peninsula, the west side tends to be relatively warmer. The range of MAT at the sites over our study period is 2-10℃. Gihwari and Gigok sites have the lowest (3℃) and Jeungsan-ri site has the highest MAT (8℃). The mean MAT for all sites fluctuated within 4℃ between 50 ka and 10 ka. The first appearance of stemmed points occurred in the middle of the decreasing MAT trend in MIS 3, at 40-35 ka. Figure \@ref(fig:climate-model)D shows a negative relationship between temperature and elevation. For example, Gihwari Cave site, with one of the lowest MAT distributions, is located in the highest elevation around 300m above sea level. Jeungsanri is located at a lower elevation and has one of the highest MAT distributions.
